\documentclass[a4paper]{article}
\usepackage{fullpage}
\usepackage{graphicx}
\usepackage{xcolor}

\IfFileExists{.git/gitHeadInfo.gin}{
    \usepackage[pcount,grumpy,mark,markifdirty]{gitinfo2}
}{%
    \usepackage[local,pcount,grumpy,mark,markifdirty]{gitinfo2}
}



\input{preamble}
\usepackage{hyperref}
\usepackage[noabbrev]{cleveref}
\usepackage{xurl}

\makeatletter
\title{Model order reduction and sensitivity analysis for complex heat transfer simulations inside the human eyeball}
\author{%
    \texorpdfstring{
        Thomas Saigre\textsuperscript{1}\textsuperscript{$\dagger$}, Christophe Prud'homme\textsuperscript{1}, Marcela Szopos\textsuperscript{2}\\
        \small\textsuperscript{1}Institut de Recherche Mathématique Avancée, UMR 7501 Université de Strasbourg et CNRS\\
        \small\textsuperscript{2}Université Paris Cité, CNRS, MAP5, F-75006 Paris, France\\
        \small\textsuperscript{$\dagger$}Corresponding author contact: \texttt{saigre@math.unistra.fr}
    }{Thomas Saigre, Christophe Prud'homme, Marcela Szopos}
}
\date{\gitReln\  \gitAuthorDate\ (\gitAbbrevHash)}

% Define custom color
\definecolor{CustomBlue}{rgb}{0.25, 0.41, 0.88} % RoyalBlue
% Set up hyperref with the custom citecolor
\hypersetup{
    pdftitle={\@title},
    pdfauthor={\@author},
    % pdfsubject={\@subject},
    pdfkeywords={LaTeX, GitHub Actions, Zotero, Overleaf},
    bookmarksnumbered,bookmarksopen,linktocpage,
    colorlinks=true,
    citecolor=CustomBlue,
    linkcolor=CustomBlue,
    urlcolor=blue
}
\makeatother
\renewcommand{\d}{\,\mathrm{d}}


\begin{document}

\maketitle

\begin{abstract}
    Heat transfer in the human eyeball, a complex organ, is significantly influenced by various pathophysiological and external parameters.
    Particularly, heat transfer critically affects fluid behavior within the eye and ocular drug delivery processes.
    Overcoming the challenges of experimental analysis, this study introduces a comprehensive three-dimensional mathematical and computational model to simulate the heat transfer in a realistic geometry.
    Our work includes an extensive sensitivity analysis to address uncertainties and delineate the impact of different variables on heat distribution in ocular tissues.
    To manage the model's complexity, we employed a very fast model reduction technique with certified sharp error bounds, ensuring computational efficiency without compromising accuracy.
    Our results demonstrate remarkable consistency with experimental observations and align closely with existing numerical findings in the literature.
    Crucially, our findings underscore the significant role of blood flow and environmental conditions, particularly in the eye's internal tissues.
    Clinically, this model offers a promising tool for examining the temperature-related effects of various therapeutic interventions on the eye.
    Such insights are invaluable for optimizing treatment strategies in ophthalmology.
\end{abstract}

\tableofcontents



\vspace{\baselineskip}

\noindent
\textbf{Keywords :} mathematical and computational ophthalmology, heat transfer, validation, finite element method, real-time model order reduction, uncertainty quantification, sensitivity analysis, Sobol index analysis.

% 1. Introduction
\input{tex/1-intro}

% 2. Three-dimensional biophysical model
\input{tex/2-model-phys}

% 3. Mathematical and computational framework
\input{tex/3-model-num}

% 4. Uncertainty quantificationœ
\input{tex/4-uq}

% 5. Conclusion
\input{tex/5-conclusion}

\appendix

\input{tex/A-reproduce}


\section*{Acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

The authors would like to acknowledge the support of the platform Cemosis at University of Strasbourg and the French Ministry of Higher Education, Research and Innovation.
This work has benefited from a national grant managed by the French National Research Agency (Agence Nationale de la Recherche) attributed to the Exa-MA project of the NumPEx PEPR program, under the reference ANR-22-EXNU-0002.


\printbibliography{}
\addcontentsline{toc}{section}{References}



\end{document}